### 2009-12-09 trunk/Papers/reasoning-challenges/conclusion.tex
ICAPS submission for Reasoning Challenges 2.0

In this paper, we introduced two novel planning domains arising from
problems in natural language generation: the sentence generation
domain and the GIVE navigation domain. We also reported on the results
of a number of experiments in which we applied off-the-shelf planners
to a set of problem instances in these domains.

Our results were mixed. While modern planners do a pretty good job of
controlling the complexity of search, they also suffer from practical
problems that limit their performance in unexpected ways. In particular, in
both domains the grounding step performed by FF and SGPLAN dominates the
time it takes to perform the search itself. In future work, we hope to
extend our analysis by exploring other planners, such as SHOP2
\cite{DBLP:journals/jair/NauAIKMWY03}, which do not perform the initial
grounding step. We also believe our domains would provide suitable
challenges for planners entered in future editions of the IPC.
%We are also considering to propose some of our domains as IPC benchmarks.

The NLG community's recent interest in planning presents a valuable
opportunity for planning researchers.  While GIVE highlights
certain challenges such as plan execution monitoring and plan presentation
(i.e., summarisation and elaboration), it also offers a platform on which
such technologies can be evaluated in experiments with human users.
Furthermore, research related to reasoning under uncertainty, resource
management, and planning with knowledge and sensing, can also be
investigated in this domain setting. Provided some of the challenges we
have highlighted can be addressed, projects like GIVE (and other
NLG-inspired problem domains) offer a constructive platform for the
planning community to showcase their techniques to a wider audience---and
to improve the quality of their tools for real-world planning tasks.

%The NLG community's recent interest in planning presents a valuable
%opportunity for planning researchers: provided some of the challenges
%we have highlighted can be addressed, projects like GIVE (and other
%NLG-inspired problem domains) offer a constructive platform for the
%planning community to showcase their techniques to a wider
%audience---and to improve the quality of their tools for real-world
%planning tasks.  Furthermore, GIVE highlights planning-related
%challenges such as plan execution monitoring and plan presentation
%and summarisation and offers a platform on which such technologies can
%be evaluated in experiments with human users.

\section{Acknowledgements}
This work arose in the context of the Planning and Language Interest Group
at the University of Edinburgh. We thank all members of this group,
especially Hector Geffner and Mark Steedman, for interesting discussions.
This work was partially supported by the DFG Research Fellowship ``CRISP:
Efficient integrated realization and microplanning'' and by the European
Commission through the PACO-PLUS project (FP6-2004-IST-4-27657).

%\paragraph{Acknowledgments.} This work is a collaboration that arose
%in the context of the ``planning and language'' group at the
%University of Edinburgh.  We thank all members of this group,
%especially Hector Geffner and Mark Steedman, for interesting
%discussions.  This work was supported by the DFG Research Fellowship
%``CRISP: Efficient integrated realization and microplanning'' and the
%EU project XYZ PACO-PLUS.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "experiences"
%%% End:

>>>>

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "main"
%%% End:

### 2009-12-09 trunk/Papers/reasoning-challenges/introduction.tex
ICAPS submission for Reasoning Challenges 2.0

The generation of referring expressions (GRE) is one of the most
active and successful research areas in natural language generation.
Building upon Dale and Reiter's work \cite{dale89cooking,Dale1995},
various researchers have added extensions such as reference to sets
\cite{Stone2000}, more expressive logical connectives
\cite{deemter02:_gener_refer_expres}, and relational expressions
\cite{dale91:_gener_refer_expres_invol_relat}.

Referring expressions (REs) involving relations, in particular, have
received increasing attention recently; especially in the context of
spatial referring expressions in situated generation
\cite<e.g.,>{kelleher06:_increm_gener_of_spatial_refer}, where it seems
particularly natural to use expressions such as ``the book on the
table''.  However, the classical algorithm by
\citeA{dale91:_gener_refer_expres_invol_relat} was recently shown to
be unable to generate satisfying REs in practice
\cite{viethen06:_algor_for_gener_refer_expres}.  Furthermore, the Dale
and Haddock algorithm and most of its successors \cite<such
as>{kelleher06:_increm_gener_of_spatial_refer} are vulnerable to the
problem of ``infinite regress'', where the algorithm jumps back and
forth between generating descriptions for two related individuals
infinitely, as in ``the book on the table which supports a book on the
table \ldots''.

In this paper, we propose to view GRE as the problem of computing a
formula of description logic (DL) that denotes exactly the set of
individuals that we want to refer to.  This very natural idea has been
mentioned in passing before
\cite{Krahmer2003,gardent07:_gener_bridg_defin_descr}; however, we
take it one step further by proposing DL as an interlingua for
comparing the REs produced by different approaches to GRE.  In this
way, we can organize existing GRE approaches in an expressiveness
hierarchy.  For instance, the classical Dale and Reiter algorithms
compute purely conjunctive formulas;
\citeA{deemter02:_gener_refer_expres} extends this language by
adding the other propositional connectives, whereas
\citeA{dale91:_gener_refer_expres_invol_relat} extends it by
allowing existential quantification.

Furthermore, the view of GRE as a problem of computing DL formulas
with a given extension allows us to apply existing algorithms for the
latter problem to obtain efficient algorithms for GRE.  We present
algorithms that compute such formulas for the description logics \el\
(which allows only conjunction and existential quantification) and
\alc\ (which also allows negation).  These algorithms effectively
compute REs for all individuals in the domain at the same time, which
allows them to systematically avoid the infinite regress problem.  The
\el\ algorithm is capable of generating 67\% of the relational REs in
the \citeA{viethen06:_algor_for_gener_refer_expres} dataset, in about
15 milliseconds.  The \alc\ algorithm is even faster; it computes
relational REs for all 100 individuals in a random model in 140
milliseconds.

The paper is structured as follows.  In Section~\ref{sec:bisim}, we
will first define description logics.  We will then show how to
generate REs by computing DL similarity sets for \alc\ and \el\ in
Section~\ref{sec:gre}.  In Section~\ref{sec:discussion}, we evaluate
our algorithms and discuss our results.  Section~\ref{sec:related}
compares our approach to related research; in particular, it shows how
various prominent GRE algorithms fit into the DL framework.
Section~\ref{sec:conclusion} concludes and points to future work.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "dl-gre-08"
%%% End:

>>>>

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "main"
%%% End:
